### 2025-04-17 Manuscript_Assignment.qmd
commit

---
title: "Manuscript Assignment = Calcasieu"
author: "Nicolas Rivas"
format: 
 html:
    code-fold: true
    code-summary: "Show code"
    toc: true            # Adds a table of contents
    toc-depth: 3         # Sets TOC depth to 3 levels
    number-sections: true # Enables numbered sections
editor: visual
---

```{r Load Libraries}
#| echo: FALSE
#| include: FALSE

library(tidyverse)
library(hypervolume)
library(MuMIn)

```

```{r load data}
#| echo: TRUE
#| include: FALSE

df = read.csv("data/LDWFseinedata.csv") |> 
  filter(basin == "Calcasieu") |> 
  rename_with(tolower) |> 
  mutate(across(where(is.character), tolower),
         year = substr(date, 1,4),
         month = substr(date, 5,6),
         day = substr(date, 7,8)) |>  
  rename(turbidity = tubidity) |> 
  select(-date)

df2 = readRDS('data/first_hvs.rds') |> 
  mutate(year=as.numeric(year))

df_ov=read_csv("data/first_centDist.csv")

dfs_hvs = readRDS('data/first_hvs_species.rds')

suitability_df = readRDS('data/suitability_df.rds')
```

# Create HVs

```{r create HVs}
#| eval: FALSE
#| echo: TRUE

set.seed(1)

## Scaling ####
df2 <- df |> 
  group_by(lat, lon, year, month, day) |> 
  slice(1) |> 
  ungroup() |>
  mutate(across(salinity:airtemp, scale)) |> 
  select(year,salinity, watertemp, turbidity,airtemp) |> 
  group_by(year) |> 
  nest()

head(df2)

## HVs Creation ####

df2 = df2 |> 
  mutate(hv = map(data, \(data) hypervolume_gaussian(data, name = paste(year),
                                                     samples.per.point = 1000,
                                                     kde.bandwidth = estimate_bandwidth(data), 
                                                     sd.count = 3, 
                                                     quantile.requested = 0.95, 
                                                     quantile.requested.type = "probability", 
                                                     chunk.size = 1000, 
                                                     verbose = F)),
         centroid = map(hv, \(hv) get_centroid(hv)),
         size = map_dbl(hv, \(hv) get_volume(hv)))

#saveRDS(df2, 'data/first_hvs.rds')

```

# Create Yearly Comparisons

```{r yearly comparisons}
#| eval: FALSE
#| echo: TRUE

# comparison of across each year
df_y= tibble(y1 = unique(df2$year),
             y2 = unique(df2$year)) |> 
  expand(y1,y2)

# make all unique year comparisons 
df_y = df_y[!duplicated(t(apply(df_y,1,sort))),] %>% 
  filter(!(y1 == y2))

# make two df to join all unique comparisons  
df3 = df2 |> 
  select( y1 = year, hv1 = hv, hv1_size = size, cent1 = centroid)

df4 = df2 |> 
  select( y2 = year, hv2 = hv, hv2_size = size, cent2 = centroid)

# create data frame of all data and make yearly comparisons
df_ov = tibble(year = rep(unique(df2$year),
                           each = nrow(df_y)),
               y1 = rep(df_y$y1, times = length(unique(df2$year))),
               y2 = rep(df_y$y2, times = length(unique(df2$year)))) |> 
  inner_join(df3, by = c('y1')) |> 
  inner_join(df4, by = c('y2')) |> 
  mutate(ychange = y2-y1,
         # calculate the differnces in size 
         lsr = log(hv2_size/hv1_size),
         # join hypervolumees in a set for cverlap
         set = map2(hv1,hv2, \(hv1, hv2) hypervolume_set(hv1, hv2, check.memory = F, verbose = F)),
         # calculate overlap
         ov = map(set, \(set) hypervolume_overlap_statistics(set)),
         # calculate centroid distance 
         dist_cent = 
           map2_dbl(hv1, hv2, \(hv1,hv2) hypervolume_distance(hv1, hv2, type = 'centroid', check.memory=F))) |> 
  #unnest centroid differences
  unnest_wider(ov) |> 
  # select only metrics of interest
  select(year, y1, y2, ychange,lsr,
         dist_cent, jaccard, sorensen,
         uniq_y1 = frac_unique_1, uniq_y2 = frac_unique_2)

beepr::beep(2)

# save output
#write_csv(df_ov, 'data/first_centDist.csv')

```

# Temporal stability

## Overlap

Overlap is calculated using the Sorensen index, which is a measure of similarity between two hypervolumes. The Sorensen index ranges from 0 to 1, where 0 indicates no overlap and 1 indicates complete overlap. We see a linear negative trend across time, indicating that the overlap between hypervolumes decreases.

```{r temporal stability - overlap}
#| echo: TRUE

# overlap 
df_o = df_ov |> 
 # group_by(basin) |>
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(sorensen~1, data = df)),
         m_lin = map(data, \(df)lm(sorensen~ychange, data = df)),
         m_quad = map(data, \(df)lm(sorensen~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_o = df_o |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_o, aes(ychange, sorensen))+
  geom_point(size = 2.5)+
  geom_smooth(data = d_o |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'black')+
  geom_smooth(data = d_o |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_o |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'black')+
  labs(x = 'Years between comparison', y = 'Overlap')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

## Centroid Distance

Centroid distance increased with time, indicating that the centroid of the hypervolume shifted over time. The linear model was the best fit for the data, indicating a linear increase in centroid distance over time.

```{r temporal stability - Centroid Distance}
#| echo: TRUE

# centroid distance
df_cd = df_ov |> 
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(dist_cent~1, data = df)),
         m_lin = map(data, \(df)lm(dist_cent~ychange, data = df)),
         m_quad = map(data, \(df)lm(dist_cent~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_cd = df_cd |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_cd, aes(ychange, dist_cent))+
  geom_hline(aes(yintercept = 1), linetype = 'dashed')+
  geom_point(size = 2.5)+
  geom_smooth(data = d_cd |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_cd |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_cd |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'blue')+
  labs(x = 'Years between comparison', y = 'Centroid distance')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

## Log Size Ratio

The log size ratio (LSR) is a measure of the change in size of the hypervolume over time. A LSR of 1 indicates no change in size, while a LSR greater than 1 indicates an increase in size and a LSR less than 1 indicates a decrease in size. The LSR decreased over time, indicating that the hypervolume shrank over time. The quadratic model was the best fit for the data, indicating a non-linear decrease in size over time.

```{r temporal stability - log size ratio}
#| echo: TRUE

# log size ratio
df_lsr = df_ov |> 
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(lsr~1, data = df)),
         m_lin = map(data, \(df)lm(lsr~ychange, data = df)),
         m_quad = map(data, \(df)lm(lsr~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_lsr = df_lsr |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_lsr, aes(ychange, lsr))+
  geom_hline(aes(yintercept = 1), linetype = 'dashed')+
  geom_point(size = 2.5)+
  geom_smooth(data = d_lsr |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_lsr |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_lsr |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'red')+
  labs(x = 'Years between comparison', y = 'log(y2 size/y1 size)')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

# Species

Using white shrimp, blue crab, brown shrimp

1.  White Shrimp (Litopenaeus setiferus) Estuarine-dependent species Spawns offshore, juveniles migrate into estuaries Highly sensitive to salinity and temperature

2.  Brown Shrimp (Farfantepenaeus aztecus) — Closely related competitor Similar habitat use to white shrimp Slightly different seasonal and salinity preferences Useful for testing niche differentiation or competitive overlap

3.  Blue Crab (Callinectes sapidus) — Functionally different predator Also estuarine-dependent Omnivorous predator, unlike shrimp Occupies broader salinity/temperature range Useful for testing niche separation between trophic levels

## First filter, scale and create HVs

```{r species}
#| echo: TRUE
#| eval: FALSE

dfs_hvs <- df |>
  filter(species %in% c("white shrimp", "blue crab", "brown shrimp")) |>
  mutate(across(c(salinity, watertemp, turbidity, airtemp), scale)) |>
  select(species, salinity, watertemp, turbidity, airtemp, num) |>
  group_by(species) |>
  nest(weight = num, data = salinity:airtemp) |> 
  mutate(hv = map2(data,weight, \(data,weight) hypervolume_gaussian(data, name = paste(species),
                                                     samples.per.point = 1000,
                                                     weight = weight$num,
                                                     kde.bandwidth = estimate_bandwidth(data), 
                                                     sd.count = 3, 
                                                     quantile.requested = 0.95, 
                                                     quantile.requested.type = "probability", 
                                                     chunk.size = 1000, 
                                                     verbose = F)),
         centroid = map(hv, \(hv) get_centroid(hv)),
         size = map_dbl(hv, \(hv) get_volume(hv)))

beepr::beep(3)

```

```{r species - plot hvs}
#| echo: TRUE

hvj = hypervolume_join(dfs_hvs$hv[[1]], dfs_hvs$hv[[2]], dfs_hvs$hv[[3]])
plot(hvj,
     colors = c('#0000CC', '#D72638', '#228B22'))

```

```{r species - comparisons}
#| echo: TRUE
#| eval: FALSE

# comparison of across each species
df_y= tibble(y1 = unique(dfs_hvs$species),
             y2 = unique(dfs_hvs$species)) |> 
  expand(y1,y2)

# make all unique species comparisons 
df_y = df_y[!duplicated(t(apply(df_y,1,sort))),] %>% 
  filter(!(y1 == y2))

# make two df to join all unique comparisons  
df1 = dfs_hvs |> 
  select(y1 = species, hv1 = hv, hv1_size = size, cent1 = centroid)

df2 = dfs_hvs |> 
  select(y2 = species, hv2 = hv, hv2_size = size, cent2 = centroid)

# create data frame of all data and make species comparisons
df_ov = tibble(species = rep(unique(dfs_hvs$species),
                           each = nrow(df_y)),
               y1 = rep(df_y$y1, times = length(unique(dfs_hvs$species))),
               y2 = rep(df_y$y2, times = length(unique(dfs_hvs$species)))) |> 
  inner_join(df1, by = c( 'y1')) |> 
  inner_join(df2, by = c( 'y2')) |> 
  mutate(
         # calculate the differnces in size 
         lsr = log(hv2_size/hv1_size),
         # join hypervolumees in a set for cverlap
         set = map2(hv1,hv2, \(hv1, hv2) hypervolume_set(hv1, hv2, check.memory = F, verbose = F)),
         # calculate overlap
         ov = map(set, \(set) hypervolume_overlap_statistics(set)),
         # calculate centroid distance 
         dist_cent = map2_dbl(hv1, hv2, \(hv1,hv2) hypervolume_distance(hv1, hv2, type = 'centroid', check.memory=F))) |> 
  #unnest centroid differences
  unnest_wider(ov) |> 
  # select only metrics of interest
  select(y1, y2,lsr,
         dist_cent, jaccard, sorensen,
         uniq_y1 = frac_unique_1, uniq_y2 = frac_unique_2)

# save output
#write_csv(df_ov, 'data/species_centDist.csv')

```

# Suitability

## Load Data make comparisons

```{r suitability}
#| echo: TRUE
#| eval: FALSE
hv_species = dfs_hvs |> 
  select(species, hv)

# Keep only environmental data
env_scaled = df |>
  filter(species %in% c("white shrimp", "blue crab", "brown shrimp")) |>
  select(year, salinity, watertemp, turbidity, airtemp) |>
  drop_na() |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(salinity:airtemp, scale)) |> 
  group_by(year) |> 
  nest()

# Expand grid of species × years
suitability_df <- crossing(
  species = c("white shrimp", "blue crab", "brown shrimp"),
  year = unique(env_scaled$year)
)

# Join HVs and data
suitability_df <- suitability_df |>
  left_join(hv_species, by = "species") |>
  left_join(env_scaled, by = "year")

# Calculate % of suitable environments
suitability_df <- suitability_df |>
  mutate(prop_suitable = map2_dbl(hv, data, ~{
    points_inside <- hypervolume::hypervolume_inclusion_test(.x, .y)
    mean(points_inside)  # proportion of sites inside the niche
  }))

# Caused by warning in `hypervolume::hypervolume_inclusion_test()`:
#   ! Results may have a high error rate. Consider setting fast.or.accurate='accurate'.
```

Habitat suitability seems to remain stable for blue crabs, fluctuates for brown shrimp but generally stable and decreases for white shrimp near 2018

```{r suitability - plot}
ggplot(suitability_df, aes(x = year, y = prop_suitable, color = species)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  theme_bw() +
  labs(x = "Year", y = "Proportion of Suitable Sites",
       title = "Habitat Suitability Over Time")

```