### 2025-04-18 Manuscript_Assignment.qmd
c

---
title: "Manuscript Assignment = Calcasieu"
author: "Nicolas Rivas"
format: 
 html:
    code-fold: true
    code-summary: "Show code"
    toc: true            # Adds a table of contents
    toc-depth: 3         # Sets TOC depth to 3 levels
    number-sections: true # Enables numbered sections
editor: visual
---

```{r Load Libraries}
#| echo: FALSE
#| include: FALSE

library(tidyverse)
library(hypervolume)
library(MuMIn)
library(ggpubr)
library(gt)

```

```{r load data}
#| echo: TRUE
#| include: FALSE

df = read.csv("data/LDWFseinedata.csv") |> 
  filter(basin == "Calcasieu") |> 
  rename_with(tolower) |> 
  mutate(across(where(is.character), tolower),
         year = substr(date, 1,4),
         month = substr(date, 5,6),
         day = substr(date, 7,8)) |>  
  rename(turbidity = tubidity) |> 
  select(-date)

df2 = readRDS('data/first_hvs.rds') |> 
  mutate(year=as.numeric(year))

df_ov=read_csv("data/first_centDist.csv")

dfs_hvs = readRDS('data/first_hvs_species.rds')

suitability_df = readRDS('data/suitability_df.rds')
```

# Create HVs

```{r create HVs}
#| eval: FALSE
#| echo: TRUE

set.seed(1)

## Scaling ####
df2 <- df |> 
  group_by(lat, lon, year, month, day) |> 
  slice(1) |> 
  ungroup() |>
  mutate(across(salinity:airtemp, scale)) |> 
  select(year,salinity, watertemp, turbidity,airtemp) |> 
  group_by(year) |> 
  nest()

head(df2)

## HVs Creation ####

df2 = df2 |> 
  mutate(hv = map(data, \(data) hypervolume_gaussian(data, name = paste(year),
                                                     samples.per.point = 1000,
                                                     kde.bandwidth = estimate_bandwidth(data), 
                                                     sd.count = 3, 
                                                     quantile.requested = 0.95, 
                                                     quantile.requested.type = "probability", 
                                                     chunk.size = 1000, 
                                                     verbose = F)),
         centroid = map(hv, \(hv) get_centroid(hv)),
         size = map_dbl(hv, \(hv) get_volume(hv)))

#saveRDS(df2, 'data/first_hvs.rds')

```

# Create Yearly Comparisons

```{r yearly comparisons}
#| eval: FALSE
#| echo: TRUE

# comparison of across each year
df_y= tibble(y1 = unique(df2$year),
             y2 = unique(df2$year)) |> 
  expand(y1,y2)

# make all unique year comparisons 
df_y = df_y[!duplicated(t(apply(df_y,1,sort))),] %>% 
  filter(!(y1 == y2))

# make two df to join all unique comparisons  
df3 = df2 |> 
  select( y1 = year, hv1 = hv, hv1_size = size, cent1 = centroid)

df4 = df2 |> 
  select( y2 = year, hv2 = hv, hv2_size = size, cent2 = centroid)

# create data frame of all data and make yearly comparisons
df_ov = tibble(year = rep(unique(df2$year),
                           each = nrow(df_y)),
               y1 = rep(df_y$y1, times = length(unique(df2$year))),
               y2 = rep(df_y$y2, times = length(unique(df2$year)))) |> 
  inner_join(df3, by = c('y1')) |> 
  inner_join(df4, by = c('y2')) |> 
  mutate(ychange = y2-y1,
         # calculate the differnces in size 
         lsr = log(hv2_size/hv1_size),
         # join hypervolumees in a set for cverlap
         set = map2(hv1,hv2, \(hv1, hv2) hypervolume_set(hv1, hv2, check.memory = F, verbose = F)),
         # calculate overlap
         ov = map(set, \(set) hypervolume_overlap_statistics(set)),
         # calculate centroid distance 
         dist_cent = 
           map2_dbl(hv1, hv2, \(hv1,hv2) hypervolume_distance(hv1, hv2, type = 'centroid', check.memory=F))) |> 
  #unnest centroid differences
  unnest_wider(ov) |> 
  # select only metrics of interest
  select(year, y1, y2, ychange,lsr,
         dist_cent, jaccard, sorensen,
         uniq_y1 = frac_unique_1, uniq_y2 = frac_unique_2)

beepr::beep(2)

# save output
#write_csv(df_ov, 'data/first_centDist.csv')

```

Overlap
```{r yearly comparisons - plot}
#| echo: FALSE
d = df_ov |> 
  filter(ychange == 1)

# overlap
ggplot(d, aes(y2, sorensen,color="blue"))+
  geom_point(size = 2)+
  geom_line(linewidth = 1)+
  labs(x = 'Year', y = 'Overlap')+
  scale_fill_viridis_d(option = 'turbo')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  #facet_wrap(~BASIN, nrow = 2)+
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 14, colour = "gray0"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

Centroid Distance
```{r yearly comparisons - plot2}
#| echo: FALSE

# centroid distance 
ggplot(d, aes(y2, dist_cent))+
  geom_hline(aes(yintercept = 1), linetype = 'dashed', linewidth = 1)+
  geom_point(size = 2)+
  geom_line(linewidth = 1)+
  labs(x = 'Year', y = 'Centroid distance')+
  scale_fill_viridis_d(option = 'turbo')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  #facet_wrap(~BASIN, nrow = 2)+
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 14, colour = "gray0"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

LSR

```{r yearly comparisons - plot3}
#| echo: FALSE

# log size ratio
ggplot(d, aes(y2, lsr))+
  geom_hline(aes(yintercept = 0), linetype = 'dashed', linewidth = 1)+
  geom_point(size = 2)+
  geom_line(linewidth = 1)+
  labs(x = 'Year', y = 'log(y2 size/y1 size)')+
  scale_fill_viridis_d(option = 'turbo')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  #facet_wrap(~BASIN, nrow = 2)+
  theme(axis.title = element_text(size = 14), 
        axis.text = element_text(size = 14, colour = "gray0"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

# Temporal stability

## Overlap

Overlap is calculated using the Sorensen index, which is a measure of similarity between two hypervolumes. The Sorensen index ranges from 0 to 1, where 0 indicates no overlap and 1 indicates complete overlap. We see a linear negative trend across time, indicating that the overlap between hypervolumes decreases.

```{r temporal stability - overlap}
#| echo: TRUE

# overlap 
df_o = df_ov |> 
 # group_by(basin) |>
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(sorensen~1, data = df)),
         m_lin = map(data, \(df)lm(sorensen~ychange, data = df)),
         m_quad = map(data, \(df)lm(sorensen~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_o = df_o |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_o, aes(ychange, sorensen))+
  geom_point(size = 2.5)+
  geom_smooth(data = d_o |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'black')+
  geom_smooth(data = d_o |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_o |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'black')+
  labs(x = 'Years between comparison', y = 'Overlap')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

## Centroid Distance

Centroid distance increased with time, indicating that the centroid of the hypervolume shifted over time. The linear model was the best fit for the data, indicating a linear increase in centroid distance over time.

```{r temporal stability - Centroid Distance}
#| echo: TRUE

# centroid distance
df_cd = df_ov |> 
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(dist_cent~1, data = df)),
         m_lin = map(data, \(df)lm(dist_cent~ychange, data = df)),
         m_quad = map(data, \(df)lm(dist_cent~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_cd = df_cd |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_cd, aes(ychange, dist_cent))+
  geom_hline(aes(yintercept = 1), linetype = 'dashed')+
  geom_point(size = 2.5)+
  geom_smooth(data = d_cd |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_cd |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_cd |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'blue')+
  labs(x = 'Years between comparison', y = 'Centroid distance')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

## Log Size Ratio

The log size ratio (LSR) is a measure of the change in size of the hypervolume over time. A LSR of 1 indicates no change in size, while a LSR greater than 1 indicates an increase in size and a LSR less than 1 indicates a decrease in size. The LSR decreased over time, indicating that the hypervolume shrank over time. The quadratic model was the best fit for the data, indicating a non-linear decrease in size over time.

```{r temporal stability - log size ratio}
#| echo: TRUE

# log size ratio
df_lsr = df_ov |> 
  nest() |> 
  # fit intercept, linear, and quadratic model
  mutate(m_int = map(data, \(df)lm(lsr~1, data = df)),
         m_lin = map(data, \(df)lm(lsr~ychange, data = df)),
         m_quad = map(data, \(df)lm(lsr~ychange + I(ychange^2), data = df)),
         AICc_int = map_dbl(m_int, \(x) AICc(x)),
         AICc_lin = map_dbl(m_lin, \(x) AICc(x)),
         AICc_quad = map_dbl(m_quad, \(x) AICc(x)),
         model = case_when(
           AICc_int - min(c(AICc_int,AICc_lin,AICc_quad)) <= 4 ~ 'Intercept',
           AICc_lin - AICc_quad <= 4 ~ 'Linear',
           T ~ 'Quadratic'))

# unnest data 
d_lsr = df_lsr |> 
  select(data, model) |> 
  unnest(cols = c(data))

ggplot(d_lsr, aes(ychange, lsr))+
  geom_hline(aes(yintercept = 1), linetype = 'dashed')+
  geom_point(size = 2.5)+
  geom_smooth(data = d_lsr |> filter(model == 'Intercept'),
              method = 'lm', formula = y~1, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_lsr |> filter(model == 'Linear'),
              method = 'lm', formula = y~x, 
              linewidth = 1, color = 'blue')+
  geom_smooth(data = d_lsr |> filter(model == 'Quadratic'),
              method = 'lm', formula = y~x+I(x^2), 
              linewidth = 1, color = 'red')+
  labs(x = 'Years between comparison', y = 'log(y2 size/y1 size)')+
  scale_color_viridis_d(option = 'turbo')+
  theme_bw()+
  theme(axis.title = element_text(size = 14), 
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        plot.title = element_text(size = 14, hjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none',
        legend.title = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        legend.text = element_text(size = 12))

```

# Species

Using white shrimp, blue crab, brown shrimp

1.  White Shrimp (Litopenaeus setiferus) Estuarine-dependent species Spawns offshore, juveniles migrate into estuaries Highly sensitive to salinity and temperature

2.  Brown Shrimp (Farfantepenaeus aztecus) — Closely related competitor Similar habitat use to white shrimp Slightly different seasonal and salinity preferences Useful for testing niche differentiation or competitive overlap

3.  Blue Crab (Callinectes sapidus) — Functionally different predator Also estuarine-dependent Omnivorous predator, unlike shrimp Occupies broader salinity/temperature range Useful for testing niche separation between trophic levels

## First filter, scale and create HVs

```{r species}
#| echo: TRUE
#| eval: FALSE

dfs_hvs <- df |>
  filter(species %in% c("white shrimp", "blue crab", "brown shrimp")) |>
  mutate(across(c(salinity, watertemp, turbidity, airtemp), scale)) |>
  select(species, salinity, watertemp, turbidity, airtemp, num) |>
  group_by(species) |>
  nest(weight = num, data = salinity:airtemp) |> 
  mutate(hv = map2(data,weight, \(data,weight) hypervolume_gaussian(data, name = paste(species),
                                                     samples.per.point = 1000,
                                                     weight = weight$num,
                                                     kde.bandwidth = estimate_bandwidth(data), 
                                                     sd.count = 3, 
                                                     quantile.requested = 0.95, 
                                                     quantile.requested.type = "probability", 
                                                     chunk.size = 1000, 
                                                     verbose = F)),
         centroid = map(hv, \(hv) get_centroid(hv)),
         size = map_dbl(hv, \(hv) get_volume(hv)))

beepr::beep(3)

```

## HV plots

```{r species - plot hvs}
#| echo: TRUE

hvj = hypervolume_join(dfs_hvs$hv[[1]], dfs_hvs$hv[[2]], dfs_hvs$hv[[3]])
plot(hvj,
     colors = c('#0000CC', '#D72638', '#228B22'))

```

```{r species - comparisons}
#| echo: TRUE
#| eval: FALSE

# comparison of across each species
df_y= tibble(y1 = unique(dfs_hvs$species),
             y2 = unique(dfs_hvs$species)) |> 
  expand(y1,y2)

# make all unique species comparisons 
df_y = df_y[!duplicated(t(apply(df_y,1,sort))),] %>% 
  filter(!(y1 == y2))

# make two df to join all unique comparisons  
df1 = dfs_hvs |> 
  select(y1 = species, hv1 = hv, hv1_size = size, cent1 = centroid)

df2 = dfs_hvs |> 
  select(y2 = species, hv2 = hv, hv2_size = size, cent2 = centroid)

# create data frame of all data and make species comparisons
df_ov = tibble(species = rep(unique(dfs_hvs$species),
                           each = nrow(df_y)),
               y1 = rep(df_y$y1, times = length(unique(dfs_hvs$species))),
               y2 = rep(df_y$y2, times = length(unique(dfs_hvs$species)))) |> 
  inner_join(df1, by = c( 'y1')) |> 
  inner_join(df2, by = c( 'y2')) |> 
  mutate(
         # calculate the differnces in size 
         lsr = log(hv2_size/hv1_size),
         # join hypervolumees in a set for cverlap
         set = map2(hv1,hv2, \(hv1, hv2) hypervolume_set(hv1, hv2, check.memory = F, verbose = F)),
         # calculate overlap
         ov = map(set, \(set) hypervolume_overlap_statistics(set)),
         # calculate centroid distance 
         dist_cent = map2_dbl(hv1, hv2, \(hv1,hv2) hypervolume_distance(hv1, hv2, type = 'centroid', check.memory=F))) |> 
  #unnest centroid differences
  unnest_wider(ov) |> 
  # select only metrics of interest
  select(y1, y2,lsr,
         dist_cent, jaccard, sorensen,
         uniq_y1 = frac_unique_1, uniq_y2 = frac_unique_2)

# save output
#write_csv(df_ov, 'data/species_centDist.csv')

```

## Niche SIze

Nice size of the three species

```{r species - plot size}
#| echo: FALSE

dfs_hvs |>
  select(species, size) |>
  ggplot(aes(x = species, y = 1, size = size, fill = species)) +
  geom_point(shape = 21, color = "black") +
  geom_text(aes(label = round(size, 2)), vjust = -3, size = 9) +  # nicely formatted labels
  scale_size_continuous(range = c(10, 40)) +
  labs(
    title = "Niche Size of Estuarine Species",
    x = "Species",
    y = NULL
  ) +
  guides(size = "none", fill = "none") +  # remove legends
  theme_classic(base_size = 14) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

## Overlap

```{r species - table overlap}
#| echo: FALSE
#| eval: FALSE

df_ov_dedup <- df_ov |> 
  group_by(y1, y2) |> 
  summarise(across(where(is.numeric), mean), .groups = "drop")

df_ov_dedup |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  gt() |>
  tab_header(
    title = "Pairwise Niche Overlap and Dissimilarity Metrics"
  ) |>
  cols_label(
    y1 = "Species 1",
    y2 = "Species 2",
    lsr = "Log Size Ratio",
    dist_cent = "Centroid Distance",
    jaccard = "Jaccard",
    sorensen = "Sorensen",
    uniq_y1 = "Unique to Species 1",
    uniq_y2 = "Unique to Species 2"
  )

```
![Pairwise Niche Overlap Table](images/niche_overlap_table.png)

Heatmap of similarity between species using Sorensen

![Niche Overlap Heatmap](sorensen_heatmap.png)

# Suitability

## Load Data make comparisons

```{r suitability}
#| echo: TRUE
#| eval: FALSE
hv_species = dfs_hvs |> 
  select(species, hv)

# Keep only environmental data
env_scaled = df |>
  filter(species %in% c("white shrimp", "blue crab", "brown shrimp")) |>
  select(year, salinity, watertemp, turbidity, airtemp) |>
  drop_na() |>
  mutate(across(everything(), as.numeric)) |>
  mutate(across(salinity:airtemp, scale)) |> 
  group_by(year) |> 
  nest()

# Expand grid of species × years
suitability_df <- crossing(
  species = c("white shrimp", "blue crab", "brown shrimp"),
  year = unique(env_scaled$year)
)

# Join HVs and data
suitability_df <- suitability_df |>
  left_join(hv_species, by = "species") |>
  left_join(env_scaled, by = "year")

# Calculate % of suitable environments
suitability_df <- suitability_df |>
  mutate(prop_suitable = map2_dbl(hv, data, ~{
    points_inside <- hypervolume::hypervolume_inclusion_test(.x, .y)
    mean(points_inside)  # proportion of sites inside the niche
  }))

# Caused by warning in `hypervolume::hypervolume_inclusion_test()`:
#   ! Results may have a high error rate. Consider setting fast.or.accurate='accurate'.
```

## Habitat Suitability/Time

Habitat suitability seems to remain stable for blue crabs, fluctuates for brown shrimp but generally stable and decreases for white shrimp near 2018

```{r suitability - plot}
ggplot(suitability_df, aes(x = year, y = prop_suitable, color = species)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  theme_bw() +
  labs(x = "Year", y = "Proportion of Suitable Sites",
       title = "Habitat Suitability Over Time")

```

## Species Abundance

I went ahead and plotted abundance data. No brown shrimp or white shimp were caught in 2019 and in general there was a continous decrease in their abundance numbers.

We can see that sharp fluctuation in abundance is normal but did not reach zero before 2019. Panel B just has a closer look at the last three years

```{r - abundance}
#| echo: FALSE

target_years <- as.character(1986:2019)
target_species <- c("white shrimp", "brown shrimp", "blue crab")

# Fill in missing combinations with 0
species_counts <- df |>
  filter(species %in% target_species,
         year %in% target_years) |>
  group_by(year, species) |>
  summarise(n_caught = sum(num), .groups = "drop") |>
  complete(year = target_years, species = target_species, fill = list(n_caught = 0)) |>
  arrange(year, species)

a = ggplot(species_counts, aes(x = as.numeric(year), y = n_caught, color = species)) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Abundance of Estuarine Species (1986–2019)",
    x = "Year",
    y = "Number Caught",
    color = "Species"
  ) +
  theme_classic(base_size = 14)

species_counts_zoom <- species_counts |>
  filter(as.numeric(year) >= 2017 & as.numeric(year) <= 2019)

b = ggplot(species_counts_zoom, aes(x = as.numeric(year), y = n_caught, color = species)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = 2017:2019, labels = 2017:2019) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Abundance of Estuarine Species (2017–2019)",
    x = "Year",
    y = "Number Caught",
    color = "Species"
  ) +
  theme_classic(base_size = 14)

ggarrange(a, b,
          ncol = 2, nrow = 1,
          labels = c("A", "B"),
          common.legend = TRUE,
          legend = "bottom") +
  theme(plot.title = element_text(hjust = 0.5))

```

## 2019 Suitability

Upon Further investigation. we see that the habitat suitbability in 2019 changed outsize of the centroid of white shrimp and brown shrimp. It was not longer suitable and had greater changes in magnitude compared to blue crab.....(i think)

```{r - suitability - 2019}
#| echo: FALSE
#| message: false

suitability_2019_summary = read_csv("data/suitability_2019_summary.csv")

suitability_2019_summary |> 
  mutate(across(where(is.numeric), ~ round(., 2))) |>
  gt() |>
  tab_header(
    title = "Habitat Suitability and Niche Deviation (2019)"
  ) |>
  cols_label(
    species = "Species",
    prop_suitable = "Prop. Suitable",
    salinity = "Δ Salinity",
    watertemp = "Δ Water Temp",
    turbidity = "Δ Turbidity",
    airtemp = "Δ Air Temp"
  ) |>
  fmt_number(everything(), decimals = 2)

```